Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-15 doc/invt.tex
small changes

Consider the problem of computing
the distribution, $p^*_i$,
along a discrete quantity, $i$,
for a given system.
%
%For example, $i$ can be the energy, $E$,
%in a lattice spin model or the temperature index
%in a simulated tempering simulation\cite{
%marinari1992, *lyubartsev1992}.
%
For a continuous quantity, $z$,
%which is often the case in a (classical) molecular system,
we can discretize it
such that the integer $i$ represents
the index of a small interval, or a bin,
$(z, z + \Delta z)$.
%
%The distribution is normalized as
%$\sum_{i = 1}^n p^*_i = 1$.

@@

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-shifted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
Below we will use the notations, $f_{*i}$ and $\bav{f}$,
for an arbitrary variable, $f_i$.
%
Note that the distribution, $\pmb\rho$,
can be independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
and chosen for convenience.\footnote{Alternatively,
we may define $v_{*i}(t)$ such that
$\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
i.e.,
$$
v_{*i}(t) \to -\log\left(
  \frac{ (p^*_i/p_i) \, e^{ -u_i(t) } } 
  { \sum_{j=1}^n \rho_j \, (p^*_j/p_j) \, e^{ -u_j(t) } } 
\right).
$$
One can readily show that the two definitions are equivalent
for small $v_{*i}$.
%
In well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}

@@

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \notag
  %\label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho1}
\end{equation}
%
%We may write the updating matrix in $\phi_{ki}$'s
%as\cite{bussi2006}
%%
%\begin{equation}
%  w_{ij}
%  =
%  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
%  \lambda_k \, \phi_{ki} \, \phi_{kj}
%  .
%  \notag
%\end{equation}

@@

In analogy to Eq. \eqref{eq:ET_average},
%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
we can write the error after a period, $T$,
as a sum of two contributions,
%
\begin{equation}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T),
  \label{eq:error_tot}
\end{equation}
%
where the residual error
(from the decay of the initial error)
is given by
%
\begin{equation}
  \Err_R(T)
  =
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
and the asymptotic error
is given by
%
\begin{align}
  \Err_A(T)
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
  ,
\label{eq:error_asym}
\end{align}
%
where
%
\begin{equation}
  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  ,
  \label{eq:uk_def}
\end{equation}
%
and $\Gamma_k$ is computed from
the autocorrelation function
of the histogram fluctuation,
$\kappa_k(t) \equiv \bigl\langle {\tilde \zeta}_k(0)
\, {\tilde \zeta}_k(t) \bigr\rangle$,
as:
%
\begin{equation}
  \Gamma_k
  = \sum_{t = -\infty}^\infty \kappa_k(t)
  = \sum_{t = -\infty}^\infty
  \left\langle {\tilde \zeta}_k(0)
            \, {\tilde \zeta}_k(t) \right\rangle
  .
  \label{eq:Gamma_sum}
\end{equation}
%
%\begin{equation}
%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
%  \notag
%\label{eq:Gamma_integral}
%\end{equation}
%
Equation \eqref{eq:Gamma_sum} provides a way
of approximately modeling
a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
In Appendix \ref{sec:Gamma_measure},
we discuss a way of estimating $\Gamma_k$
from an FDS simulation under constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  &+
  \sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  \notag \\
  &= \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(Q') \, dQ'
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(Q) \, dQ
  =
  \frac{ 2 \, C } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible, then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

@@

\item \label{step:gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
    (cf. Appendix \ref{sec:Gamma_measure})
    as well as the initial bias $\epsilon_k$.

@@

We will first show
that the single-bin scheme is asymptotically optimal.
%
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence the same $\Gamma_k$'s,
but with different positive eigenvalues,
$\lambda_k$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of nonnegative numbers, $c_k$'s\footnote{This
inequality follows from the nonnegativity of
the quadratic polynomial,
$$
\int_0^T
  dt \sum_{k = 1}^{n-1} \Gamma_k \,
    \left( {\dot y}_k \, \chi - c_k \right)^2
  \equiv
  A_2 \, \chi^2 + A_1 \, \chi + A_0
  ,
$$
which must carry a non-positive discriminant:
$A_1^2 - 4 \, A_0 \, A_2 \le 0$.},
%
%
\begin{align}
&
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, {\dot y}_k^2\bigl( q(t) \bigr)
\right)
%
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k^2
\right)
%
\notag
\\
&
\qquad \qquad
\ge
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k \, {\dot y}_k \, \bigl( q(t) \bigr)
\right)^2
\notag
\\
&
\qquad \qquad
=
\left(
  \sum_{k = 1}^{n-1} \Gamma_k \, c_k
    \left[
      1 - e^{ -\lambda_k \, q(T) }
    \right]
\right)^2.
\label{eq:CSineq}
\end{align}
%
Note that the last expression %of Eq. \eqref{eq:CSineq}
depends on the curve, $q(t)$ ($0 < t < T$),
only through the endpoint value, $q(T)$,
which is fixed in the variational process.
%
Thus, the inequality sets a lower bound
for the asymptotic error $\Err(T)$
given in Eq. \eqref{eq:error_asym}.
%
The equality is achieved
if $\dot y_k\left( q(t) \right) = c_k$
for all $k > 0$ at any $t$,
up to a multiplicative factor.
%
Solving these equations yields
%
\begin{equation}
  \alpha(t) = \frac{              1             }
                   { \lambda_k \, (t + t_{k0} ) },
\label{eq:alpha_invtk}
\end{equation}
with
\begin{equation}
  t_{k0} = \frac{             T            }
                { e^{ \lambda_k q(T) } - 1 }.
\label{eq:tk0}
\end{equation}
%
\note{Derivation.
  Integrating $\dot y_k = c_k$ yields
  \begin{equation}
    y_k(t) = c_k \left(t + t_{k0} \right).
  \tag{UK}
  \label{neq:ui_solution}
  \end{equation}
  Given that $y_k(T) = 1$ and $y_k(0) = e^{-\lambda_k \, q(T)}$,
  we have
  $$
  c_k = \frac{ 1 }{ T + t_{k0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{k0} } { T + t_{k0} }
  =
  e^{ -\lambda_k \, q(T) }.
  $$
  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
  $$
  -\lambda_k \, q(T) + \lambda_k \, q(t)
  = \ln c_k + \ln \left( t + t_{k0} \right).
  $$
  Differentiating this with respect to $t$ yields
  Eq. \eqref{eq:alpha_invtk}.
}
%
Such a solution is possible only if
all nontrivial eigenvalues are equal,
%
%\begin{equation}
  $\lambda_1 = \dots = \lambda_{n-1}$.
%  \notag
%\end{equation}
%
Then, by Eq. \eqref{eq:tk0},
all $t_{k0}$'s also share the same value,
$t_0$,
such that
$y_k = (t + t_0) / (T + t_0)$,
and
$c_k = 1/(T + t_0)$
for $k > 0$.
%
The asymptotic error is then
%
\begin{align}
  \Err_A(T)
  =
  \frac{       T     }
       { (T + t_0)^2 }
  \sum_{ k = 1 }^{n-1}
    \Gamma_k
  .
\label{eq:error_asym_singlebin}
\end{align}
%
If we further assume $\lambda_1 = \cdots = \lambda_{n-1} = 1$,
%the optimal schedule given by
Eq. \eqref{eq:alpha_invtk}
recovers Eq. \eqref{eq:alpha_invt1}.
%

@@

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that ignore short-wavelength eigenmodes,
which are typically local noise.
%
Mathematically, these updating schemes
allow us to assign zero eigenvalues
to modes that are likely absent from the PMF.
%
For example,
%if we can assume
by assuming
that the PMF is sufficiently smooth
such that the eigenmodes can be limited to $k < K$,
%
we can construct an updating scheme with
$$
\lambda_0 = \cdots = \lambda_{K-1} = 1,
\mathrm{\; and \;}
\lambda_K = \cdots = \lambda_{n-1} = 0.
$$
%
Then, the updating matrix is given by
%
$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj},$$
%
and Eq. \eqref{eq:error_asym_singlebin} is modified as
%
\begin{equation}
  \Err_A(T)
  =
  \frac {       T     }
        { (T + t_0)^2 }
  \sum_{ k = 1 }^{K-1}
    \Gamma_k.
%\notag
\label{eq:error_asym_sinc}
\end{equation}
%
The optimal schedule of these schemes
remains to be the inverse-time schedule
given by Eq. \eqref{eq:alpha_invt1},
%
with
%
\begin{equation}
  t_0
  =
  \frac{        1      }
       { \alpha(t = 0) }
  =
  \frac{  2  }
       { a_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
where we have used Eq. \eqref{eq:half_alpha0}
in the second step.
%
\note{
For the single-bin updating scheme,
it recovers Eq. \eqref{eq:t0_singlebin}
under the assumption of Eq. \eqref{eq:xt2_eql}.
%
From Eq. \eqref{eq:xt2_eql}, we have
$$E(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
= \frac{1}{2} \, a_0 \, \Gamma.$$
}
%
From Eqs. \eqref{eq:error_tot},
\eqref{eq:error_res},
\eqref{eq:tk0},
and
\eqref{eq:error_asym_sinc},
we get
%
\begin{equation}
  \Err(T)
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 1 }^{ K - 1 }
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
This result recovers Eq. \eqref{eq:Emin_singlebin}
for the single-bin updating scheme:
if we set $K = n$, then\note{
  From Eq. \eqref{eq:eig_orthonormal_cols}, we have
  $\sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle =\sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle.$
  Then sum over $t$ from $-\infty$ to $\infty$.
}
%
\begin{equation}
  \sum_{k=1}^{n-1} \Gamma_k
  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  ,
  \notag
  %\label{eq:Gamma_G_sum}
\end{equation}
%
This error is roughly proportional to the inverse simulation length,
as one would expect from an equilibrium FDS simulation.
%In Appendix \ref{sec:equilerr},
%we show that the above error tends
%to the error of the ideal equilibrium FDS simulation
%in the asymptotic limit.

>>>>

Consider the problem of computing
the distribution, $p^*_i$,
along a discrete quantity, $i$,
for a given system.
%
%For example, $i$ can be the energy, $E$,
%in a lattice spin model or the temperature index
%in a simulated tempering simulation\cite{
%marinari1992, *lyubartsev1992}.
%
For a continuous quantity, $z$,
%which is often the case in a (classical) molecular system,
we can discretize it
such that the integer $i$ represents
the index of a small interval, or a bin,
$(z, z + \delta z)$.
%
%The distribution is normalized as
%$\sum_{i = 1}^n p^*_i = 1$.

@@

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-shifted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
Below we will use the notations, $f_{*i}$ and $\bav{f}$,
for an arbitrary variable, $f_i$.
%
Note that the distribution, $\pmb\rho$,
can be independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
and chosen for convenience.\footnote{Alternatively,
we may define $v_{*i}(t)$ such that
$\rho_i \, e^{-v_{*i}(t)}$ is a normalized distribution,
i.e.,
$$
v_{*i}(t) \to -\ln\left(
  \frac{ (p^*_i/p_i) \, e^{ -u_i(t) } }
  { \sum_{j=1}^n \rho_j \, (p^*_j/p_j) \, e^{ -u_j(t) } }
\right).
$$
One can readily show that the two definitions are equivalent
for small $v_{*i}$.
%
In well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}

@@

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \notag
  %\label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{0i} = \rho_i,
\label{eq:eigenmode0}
\end{equation}
%
with $\lambda_0 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode0}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 0$,
while the orthogonality demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \delta_{k0}
  .
\label{eq:ortho1}
\end{equation}
%
We may reconstruct the updating matrix in $\phi_{ki}$'s
as\cite{bussi2006}
%
\begin{equation}
  w_{ij}
  =
  \frac{1}{\rho_i} \sum_{k=0}^{n - 1}
  \lambda_k \, \phi_{ki} \, \phi_{kj}
  .
  \label{eq:w_from_phi}
\end{equation}

@@

In analogy to Eq. \eqref{eq:ET_average},
%using Eqs. \eqref{eq:yt_diffeq} and \eqref{eq:xt2_sum},
we can write the error after a period, $T$,
as a sum of two contributions,
%
\begin{equation}
  \Err(T)
  =
  \Err_R(T) + \Err_A(T),
  \label{eq:error_tot}
\end{equation}
%
where the residual error
(from the decay of the initial error)
is given by
%
\begin{equation}
  \Err_R(T)
  =
  \sum_{k = 1}^{n-1}
    \left\langle
      {\tilde v}_k^2(0)
    \right\rangle \,
    e^{ - 2 \, \lambda_k  \, q(T) },
  \label{eq:error_res}
\end{equation}
%
and the asymptotic error
is given by
%
\begin{align}
  \Err_A(T)
  &=
  \int_0^T
  \sum_{k = 1}^{n-1}
  \Gamma_k \, \dot y_k^2\bigl( q(t) \bigr) \, dt
  ,
\label{eq:error_asym}
\end{align}
%
where
%
\begin{equation}
  y_k(q') \equiv e^{\lambda_k \, [q' - q(T)]}
  ,
  \label{eq:uk_def}
\end{equation}
%
and $\Gamma_k$ is computed from
the autocorrelation function
of the histogram fluctuation,
$\bigl\langle {\tilde \zeta}_k(0)
\, {\tilde \zeta}_k(t) \bigr\rangle$,
as:
%
\begin{equation}
  \Gamma_k
  = \sum_{t = -\infty}^\infty
  \left\langle {\tilde \zeta}_k(0)
            \, {\tilde \zeta}_k(t) \right\rangle
  .
  \label{eq:Gamma_sum}
\end{equation}
%
%\begin{equation}
%$\Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt$,
%  \notag
%\label{eq:Gamma_integral}
%\end{equation}
%
Equation \eqref{eq:Gamma_sum} provides a way
of approximately modeling
a general histogram fluctuation, ${\tilde \zeta}_k(t)$,
as an equivalent white noise.
In this way,
the underlying sampling process
affects the error only through the
few numbers, $\Gamma_k$'s.
%
In Appendix \ref{sec:Gamma_measure},
we discuss a way of estimating $\Gamma_k$
from an FDS simulation under constant magnitude.
%

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  &+
  \sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  \notag \\
  &= \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(Q') \, dQ'
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(Q) \, dQ
  =
  \frac{ 2 \, C } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

@@

\item \label{step:gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
(cf. Appendix \ref{sec:Gamma_measure})
as well as the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it can overestimate
  $\epsilon_k$ for large-$k$ modes.
}

@@

We will first show
that the single-bin scheme is asymptotically optimal.
%
Consider a class of $\mathbf w$ matrices
sharing the same set of eigenvectors,
hence the same $\Gamma_k$'s,
but with different positive eigenvalues,
$\lambda_k$'s.
%
Using the Cauchy-Schwarz inequality, we have,
for any set of nonnegative numbers, $c_k$'s\footnote{This
inequality follows from the nonnegativity of
the quadratic polynomial,
$$
\int_0^T
  dt \sum_{k = 1}^{n-1} \Gamma_k \,
    \left( {\dot y}_k \, \chi - c_k \right)^2
  \equiv
  A_2 \, \chi^2 + A_1 \, \chi + A_0
  ,
$$
which must carry a non-positive discriminant:
$A_1^2 - 4 \, A_0 \, A_2 \le 0$.},
%
%
\begin{align}
&
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, {\dot y}_k^2\bigl( q(t) \bigr)
\right)
%
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k^2
\right)
%
\notag
\\
&
\qquad \qquad
\ge
\left(
  \int_0^T dt
    \sum_{k = 1}^{n-1}
      \Gamma_k \, c_k \, {\dot y}_k \, \bigl( q(t) \bigr)
\right)^2
\notag
\\
&
\qquad \qquad
=
\left(
  \sum_{k = 1}^{n-1} \Gamma_k \, c_k
    \left[
      1 - e^{ -\lambda_k \, q(T) }
    \right]
\right)^2.
\label{eq:CSineq}
\end{align}
%
Note that the last expression %of Eq. \eqref{eq:CSineq}
depends on the curve, $q(t)$ ($0 < t < T$),
only through the endpoint value, $q(T)$,
which is fixed in the variational process.
%
Thus, the inequality sets a lower bound
for the asymptotic error $\Err(T)$
given in Eq. \eqref{eq:error_asym}.
%
The equality is achieved
if $\dot y_k\left( q(t) \right) = c_k$
for all $k > 0$ at any $t$,
up to a multiplicative factor.
%
Solving these equations yields
%
\begin{equation}
  \alpha(t) = \frac{              1             }
                   { \lambda_k \, (t + t_{k0} ) },
\label{eq:alpha_invtk}
\end{equation}
with
\begin{equation}
  t_{k0} = \frac{             T            }
                { e^{ \lambda_k q(T) } - 1 }.
\label{eq:tk0}
\end{equation}
%
\note{Derivation.
  Integrating $\dot y_k = c_k$ yields
  \begin{equation}
    y_k(t) = c_k \left(t + t_{k0} \right).
  \tag{UK}
  \label{neq:ui_solution}
  \end{equation}
  Given that $y_k(T) = 1$ and $y_k(0) = e^{-\lambda_k \, q(T)}$,
  we have
  $$
  c_k = \frac{ 1 }{ T + t_{k0} },
  \quad
  \mathrm{and\;}
  \frac{ t_{k0} } { T + t_{k0} }
  =
  e^{ -\lambda_k \, q(T) }.
  $$
  Taking the logarithm of Eq. \eqref{neq:ui_solution} yields
  $$
  -\lambda_k \, q(T) + \lambda_k \, q(t)
  = \ln c_k + \ln \left( t + t_{k0} \right).
  $$
  Differentiating this with respect to $t$ yields
  Eq. \eqref{eq:alpha_invtk}.
}
%
Such a solution is possible only if
all nontrivial eigenvalues are equal,
%
%\begin{equation}
  $\lambda_1 = \dots = \lambda_{n-1}$.
%  \notag
%\end{equation}
%
Then, by Eq. \eqref{eq:tk0},
all $t_{k0}$'s also share the same value,
$t_0$,
such that
$y_k = (t + t_0) / (T + t_0)$,
and
$c_k = 1/(T + t_0)$
for $k > 0$.
%
The asymptotic error is then
%
\begin{align}
  \Err_A(T)
  =
  \frac{       T     }
       { (T + t_0)^2 }
  \sum_{ k = 1 }^{n-1}
    \Gamma_k
  =
  \frac{ \Gamma \, T     }
       { (T + t_0)^2 }
  ,
\label{eq:error_asym_singlebin}
\end{align}
%
where we have used
%
\begin{align}
  \sum_{k=1}^{n-1} \Gamma_k
  &=
  \sum_{t=-\infty}^\infty \sum_{k=1}^{n-1} \langle {\tilde v}_k(t) \, {\tilde v}_k(0) \rangle
  \notag
  \\
  &=
  \sum_{t=-\infty}^\infty \sum_{i=1}^n \rho_i \langle v_{*i}(t) \, v_{*i}(0) \rangle
  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  .
  \notag
  %\label{eq:Gamma_G_sum}
\end{align}
%
%
If we further assume $\lambda_1 = \cdots = \lambda_{n-1} = 1$,
%the optimal schedule given by
Eq. \eqref{eq:alpha_invtk}
recovers Eq. \eqref{eq:alpha_invt1}.
%

@@

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that ignore short-wavelength eigenmodes,
which are typically local noise.
%
A bandpass updating scheme is essentially
the single-bin updating scheme
supplied with a perfect filter
for the noise modes,
and zero eigenvalues are assigned to the noise modes.
%
For example,
by assuming
that the PMF is sufficiently smooth
such that the eigenmodes can be limited to $k < K$,
%
we can construct an updating scheme with
$$
\lambda_0 = \cdots = \lambda_{K-1} = 1,
\mathrm{\; and \;}
\lambda_K = \cdots = \lambda_{n-1} = 0.
$$
%
Then, the updating matrix is, according to
Eq. \eqref{eq:w_from_phi},
%
$$w_{ij} = n \sum_{k=0}^{K-1} \phi_{ki} \, \phi_{kj},$$
%
The optimal schedule of these schemes
remains to be the inverse-time schedule
given by Eq. \eqref{eq:alpha_invt1},
%
with
%
\begin{equation}
  t_0
  =
  \frac{        1      }
       { \alpha(t = 0) }
  =
  \frac{  2  }
       { a_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
where we have used Eq. \eqref{eq:half_alpha0}
in the second step.
%
\note{
For the single-bin updating scheme,
it recovers Eq. \eqref{eq:t0_singlebin}
under the assumption of Eq. \eqref{eq:xt2_eql}.
%
From Eq. \eqref{eq:xt2_eql}, we have
$$E(0) = \frac{1}{2} a_0 \sum_{k=1}^{n-1} \Gamma_k
= \frac{1}{2} \, a_0 \, \Gamma.$$
}
%
%From Eqs. \eqref{eq:error_tot},
%\eqref{eq:error_res},
%\eqref{eq:tk0},
%and
%\eqref{eq:error_asym_sinc},
%we get
The final error is given by
%
\begin{equation}
  \Err(T)
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 1 }^{ K - 1 }
    \Gamma_k
  ,
\label{eq:error_sinc}
\end{equation}
%
as a generalization of Eq. \eqref{eq:Emin_singlebin}
for the single-bin updating scheme.

Since the bandpass updating scheme is
a straightforward generalization of
the single-bin updating scheme,
we can also borrow the WL algorithm
to control the updating magnitude
in initial stages of the simulation.
%
However, the histogram fluctuation should
be limited to first few modes, $k = 1, \dots, K$,
as the bandpass updating scheme
only partially flattens the histogram
of these modes (cf. Appendix \ref{sec:hfluc}).

%
%This error is roughly proportional to the inverse simulation length,
%as one would expect from an equilibrium FDS simulation.
%In Appendix \ref{sec:equilerr},
%we show that the above error tends
%to the error of the ideal equilibrium FDS simulation
%in the asymptotic limit.
